---
title: "RAND Health Study" 
author: "Chris Holman"
date: "1/17/2022"
output: html_document
---
# Load Libraries
library(tidyverse)
library(broom)
library(scales)
library(fixest)
library(skimr)
library(ggplot2)

# Load Data

```{r}
data("rand_sample", package = "masteringmetrics")

devtools::install_github("jrnold/masteringmetrics", subdir = "masteringmetrics")  
```
  
# Explore Data

```{r}
varlist <- c("plantype", "female", "blackhisp", "age", "educper",
             "income1cpi", "hosp", "ghindx", "ghindxx", "cholest", "diastol",
             "systol", "mhi", "fam_identifier")
```
             

# Filter Data/ Create Data Subset
```{r}
rand_sample_small <- 
  rand_sample %>%
  select(one_of(varlist))
```



#Visualize Data

```{r}
rand_sample_small %>%
  group_by(plantype) %>%
  summarise(n = n()) %>%
  ggplot(
    aes(
      x = plantype,
      y = n)
  ) +
  geom_col(fill = "orange")
```
  
#Summarise Data
```{r}
rand_sample_small %>%
  group_by(plantype) %>%
  summarise(
    number_enrolled = n(),
    female_mean = mean(female),
    nonwhite_mean = mean(blackhisp, na.rm = T),
    age_mean = mean(age, na.rm = T),
    education_mean = mean(educper, na.rm = T),
    fam_income_mean = mean(income1cpi))
```
# cmd + shft + c Comment all
# na.rm = TRUE, will skip missing values

# Regression
```{r}
rand_sample_small %>%
  feols(income1cpi ~ plantype, data = .) %>%
  tidy(cluster = "fam_identifier")
```

#Data Summary of Person, Start Year, and Year
```{r}
rand_person_spend %>%
  select(
    person,
    indv_start_year,
    year
  )%>%
  arrange(
    person,
    year
  )
```

# Create variable for year of observation
```{r}
rand_person_spend <-
  rand_person_spend %>%
  mutate(
    expyear = indv_start_year + year - 1
  )

```

# Summarize data to confirm correct addition of new variable
```{r}

rand_person_spend %>%
  select(
    person,
    indv_start_year,
    year,
    expyear
  )%>%
  arrange(
    person,
    year
  )
```


# Create CPI Table to join to existing data
```{r}
cpi <- tribble(
  ~ year, ~ cpi,
  1973, 3.07,
  1974, 2.76,
  1975, 2.53,
  1976, 2.39,
  1977, 2.24,
  1978, 2.09,
  1979, 1.88,
  1980, 1.65,
  1981, 1.5,
  1982, 1.41,
  1983, 1.37,
  1984, 1.31,
  1985, 1.27)
```

# Join CPI table to existing data (Must run this code indepently before moving on)
```{r}
rand_person_spend <-
  rand_person_spend %>%
  left_join(cpi, by = c("expyear" = "year")
  )
```

# Summarize data to confirm correct addition of CPI table to existing data
```{r}
rand_person_spend %>%
  select(
    person,
    expyear,
    cpi
  ) %>%
  arrange(
    person,
    expyear)
```

# Create inflation 
```{r}
rand_person_spend <- 
  rand_person_spend %>%
  mutate(
    outpatient_inf = outsum * cpi,
    inpatient_inf = inpdol * cpi,
    total_inf = outpatient_inf + inpatient_inf
  )
```
# Summarize inflation
```{r}
rand_person_spend %>%
  group_by(
    plantype
  ) %>%
  summarise(
    count = n(),
    outpatient_mean = mean(outpatient_inf),
    outpatient_sd = sd(outpatient_inf),
    inpatient_mean = mean(inpatient_inf),
    inpatient_sd = sd(inpatient_inf),
    total_mean = mean(total_inf),
    total_sd = sd(total_inf)
  )
```

# Visualize health care speding by plan type
```{r}
rand_person_spend %>%
  group_by(
    plantype
  ) %>% 
  summarise(
    avg_spending = mean(total_inf)
  ) %>% 
  ggplot(
    aes(
      x = plantype,
      y = avg_spending
    )
  ) + 
  geom_col(fill = "darkgreen") +
  scale_y_continuous(labels = scales::dollar) +
  theme_minimal() +
  labs(
    y = "Average Spending ($)",
    x = "",
    title = "Average Health Expenditure by Health Plan"
  )
```

# Visualize average spending on health care by plan type from 1974-85
```{r}
rand_person_spend %>% 
  group_by(
    plantype,
    expyear
  ) %>%
  summarise(
    avg_spending = mean(total_inf)
  ) %>% 
  ggplot(
    aes(
      x = expyear,
      y = avg_spending,
      color = plantype
    )
  ) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(
    y = "Average Spending ($)",
    x = "Year",
    title = "Average Health Expenditure by Health Plan (1974-85)"
  )
```


