---
title: "Wage" 
author: "Chris Holman"
date: "1/17/2022"
output: html_document
---
#### Install Packages
```{r}
install.packages("wooldridge")

install.packages("modelsummary")
```

# Open Libraries
```{r}
library(tidyverse)
library(broom)
library(scales)
library(fixest)
library(skimr)
library(modelsummary)
library(wooldridge)
```

# Call in data
```{r}
data("wage1")
head(wage1)
?wage1
```


# Estimating gender pay gap
# OLS
```{r}
wage_reg0 <- 
  feols(
    wage ~ female,
    data = wage1
  )

summary(wage_reg0)
```
# Control for experience
```{r}
wage_reg1 <- 
  feols(
    wage ~ educ + exper + tenure + female,
    data = wage1
  )

summary(wage_reg1)
```



# Create profiles to predict wages
```{r}
new_df_male <- 
  data.frame(
  educ = 14,
  exper  = 3,
  tenure = 2,
  female = 0
)

new_df_female <- 
  data.frame(
    educ = 14,
    exper  = 3,
    tenure = 2,
    female = 1  
  )
```



# Predict wages
```{r}
wage_male1 <- predict(wage_reg1, new_df_male)
wage_female1 <- predict(wage_reg1, new_df_female)
wage_female1 - wage_male1
summary(wage_reg1)
```



# Create Table
```{r}
modelsummary(
  list(
  wage_reg0,
  wage_reg1)
)
```

# Create variables for married males, married females, and single females. (single males set as control)
```{r}
wage2 <-
  wage1 %>%
  mutate(
    marrmale = (1-female) * married,
    marrfem = female * married,
    singfem = female * (1-married)
  )
```

# Create subset dataset
```{r}
wage2 %>% 
  select (
    female,
    married,
    marrmale,
    marrfem,
    singfem
  )

```

# Regress log wage on education, experience, gender, and marital status
```{r}
wage_reg2 <-
  feols(
    log(wage) ~ marrmale + 
      marrfem + 
      singfem + 
      educ + 
      exper + 
      exper^2 + 
      tenure + 
      tenure^2,
    data = wage2
  )

summary(wage_reg2)
```
### Create Table Output
```{r}
modelsummary(
  list(
    wage_reg0,
    wage_reg1
  ),
  estimate = "{estimate}{stars} <br> ({std.error})",
  statistic = NULL,
  coef_map = c(
    "educ"= "Education",
    "female" = "Female",
    "exper" = "Experience",
    "tenure" = "Tenure"
  ),
  gof_omit = "R2 Within|R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors"
)
```
